---
title: "Homework 5"
output: html_document
---
```{r}
Crime<-read.csv("crimedata.csv")
Crime2<-read.csv("crimedata10.csv")
library(rpart)
library(rpart.plot)
summary(Crime)
```

## 2. Create Regression Tree

```{r}
CrimeTree<- rpart(Crime$CrimeRate ~Crime$ExpenditureYear+Crime$StateSize+Crime$BelowWage+Crime$HighYouthUnemploy+Crime$MatureUnemployment+Crime$Education, data = Crime)
```

## 3. Tree Summary

Per capita expendature on police and the number of low income families are the most important variables.

```{r}
summary(CrimeTree)
```


## 4. Plot the Tree

```{r}
rpart.plot(CrimeTree, digits = 3, fallen.leaves = TRUE,tweak=1.3)
```


This diagram illustrated that expenditure on police per year is the best predictor for which to divide crime rate (crime rates within groups have maximum homogenaity). The darker blue leaves indicate more certain predictions which also indicates that Expendature year is a better predictor that StateSizer, even at the second level of division. 

## 5. Mean crime rates for each group

Police Expendatures < 77 million per year: 84.5
Police Expendatures < 23 million per year: 72.5
Police Expendatures between 23 and 77 million: 97.6
Police Expendatures < 77 million: 120
Police expendatures > 108 million: 111
Police expendatures between 77 and 108 million: 131
Mean of all data: 103

## 6. Dropped variables

BelowWag, HighYouthUnemploy, MatureUnemployment and Education were all dropped from this analysis.

Minimize cross validation error

## 7. Predicting Crime rates 10 years in the future

```{r}
PredCrime <- predict(CrimeTree, Crime2)
```

## 8. Correlation of predicted and actual crime rate

```{r}
cor(PredCrime, Crime2["CrimeRate"],method="pearson")
```
The correlation coefficient is ~0.73 for this analysis.


## 9. Mean absolute error

